---
title: "Alpha_Div_WHO"
output: html_document
date: "2023-06-29"
---

MS: Figure Fungal Alpha Diversity 

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(VennDiagram)
library(car)
#library(Hmisc)
```


```{r}
otus <- read.table("DroughtITS.final_fix.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP4: Import taxonomy table for analysis using Phyloseq
This taxonomy table was created from STEP2.
```{r}
taxmat <- read.table("DroughtITS.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP2: Import tree file for analysis using Phyloseq

```{r warning=FALSE}
treefile = "DroughtITS.tree.phy"
tree = read.tree(treefile)
```

###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r}
meta = read.table("DroughtITS.mapping_file.fix.final.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

Check if your metadata file has been import successfully and correctly, the output will show a table of your metadata file (mapping file). *If you do not have header, you might start your first row with # (remove # and reload your mapping file).

```{r}
head(meta)
```

Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```


###STEP4: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
physeq
```

```{r}
physeq = subset_taxa(physeq, Kingdom == "Fungi")
#physeq = subset_samples(physeq)

```

```{r}

physeq <- subset_samples(physeq, plant == "Rice")
#physeq <- subset_samples(physeq, plant == "Cassava")
#physeq <- subset_samples(physeq, plant == "Rubber tree")
#physeq <- subset_samples(physeq, plant == "Sugarcane")

physeq

```


```{r}
physeq.two <- subset_samples(physeq, zone == "Phu_Wiang")
physeq.five <- subset_samples(physeq, zone == "Udon_Sakon_Nakon")
physeq.six <- subset_samples(physeq, zone == "Bueng_Kan")
physeq = merge_phyloseq(physeq.two, physeq.five, physeq.six)
physeq
```

```{r}
#critical group #non of the critical fungi was detected
#physeq.1 = subset_taxa(physeq, Species == "Cryptococcus neoformans")
#physeq.2 = subset_taxa(physeq, Species == "Candida auris")
#physeq.3 = subset_taxa(physeq, Species == "Aspergillus fumigatus")
#physeq.4 = subset_taxa(physeq, Species == "Candida albican")

#High group #only candida tropicalis
#physeq.5 = subset_taxa(physeq, Species == "Nakaseomyces glabrata")
#physeq.6 = subset_taxa(physeq, Genus == "Histoplasma")
physeq.7 = subset_taxa(physeq, Species == "Candida tropicalis")
#physeq.8 = subset_taxa(physeq, Species == "Candida parapsilosis")

#physeq.17 = subset_taxa(physeq, Genus == "Madurella")
physeq.18 = subset_taxa(physeq, Species == "Falciformispora senegalensis")
physeq.19 = subset_taxa(physeq, Species == "Curvularia lunata")
#physeq.20 = subset_taxa(physeq, Species == "Zopfia rosatii")
physeq.21 = subset_taxa(physeq, Genus == "Acremonium")
physeq.22 = subset_taxa(physeq, Genus == "Fusarium")
physeq.23 = subset_taxa(physeq, Genus == "Rhizopus")
physeq.24 = subset_taxa(physeq, Genus == "Mucor")
physeq.25 = subset_taxa(physeq, Genus == "Lichtheimia")

#Medium group
physeq.9 = subset_taxa(physeq, Genus == "Scedosporium")
#physeq.10 = subset_taxa(physeq, Species == "Lomentospora prolificans")
#physeq.11 = subset_taxa(physeq, Genus == "Coccidioides")
#physeq.12 = subset_taxa(physeq, Species == "Pichia kudriavzeveii")
#physeq.13 = subset_taxa(physeq, Species == "Cryptococcus gattii")
physeq.14 = subset_taxa(physeq, Species == "Talaromyces marneffei")
#physeq.15 = subset_taxa(physeq, Species == "Pneumocystis jirovecii")
#physeq.16 = subset_taxa(physeq, Genus == "Paracoccidioides")


# High group: Candida tropicalis and Medium group: Falciformispora senegalensis, Curvularia lunata, Acremonium spp, Fusarium spp, Rhizopus spp, Mucor spp, Lichtheimia, Scedosporium, Talaromyces marneffei

physeq.merge = merge_phyloseq(physeq.7, physeq.9, physeq.14, physeq.18, physeq.19, physeq.21, physeq.22, physeq.23, physeq.24, physeq.25)

physeq.merge
```

```{r}

sample_data(physeq.merge)$zone = factor(sample_data(physeq.merge)$zone, levels = c('Leoi_Petchabun', 'Phu_Wiang', 'Khorat_Plateau', 'Phu_Phan', 'Udon_Sakon_Nakon', 'Bueng_Kan', 'Khorat_Ubon' ,'Buriram')) 

```

```{r}
sample_data(physeq.merge)$zone = factor(sample_data(physeq.merge)$zone, levels = c('Phu_Wiang', 'Udon_Sakon_Nakon', 'Bueng_Kan')) 
```


```{r}
sample_data(physeq.merge)$plant <- factor(sample_data(physeq.merge)$plant, levels = c('Rice', 'Cassava', 'Rubber tree', 'Sugarcane')) 
```

```{r}
alpha.diversity = estimate_richness(physeq.merge, measures = c("Observed"))
data.anova = cbind(sample_data(physeq.merge), alpha.diversity)
physeq.anova = aov(Observed ~ zone, data.anova)
summary(physeq.anova)
```

```{r}
tukey.ps = TukeyHSD(x=physeq.anova, 'zone', conf.level = 0.95)

generate_label_df <- function(tukey.ps, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- tukey.ps[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }

LABELS=generate_label_df(tukey.ps, 'zone')
names(LABELS) = c('Letters','zone')
```

```{r}
LABELS
```

```{r}

ylabel <- data.frame("ylabel" = c(10,10,10))

```

```{r}
LABELS$ylabel<-ylabel$ylabel
```



#Alpha diversity by zone

```{r}
physeq.plot.richness.plant = plot_richness(physeq.merge, x = "zone", color=("zone"), measures=c("Observed")) + 
  geom_boxplot(lwd=0.5) + ggtitle("WHO Fungal Priority Pathogens Alpha Diversity in Rice fields by zone") + 
  stat_compare_means(method = "anova", label.y = 12) + 
  theme_pubr(border= TRUE, legend = c("none")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(l = 58) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(data=LABELS, aes(x=LABELS$zone, y = LABELS$ylabel , label = LABELS$Letters))

physeq.plot.richness.plant
```


```{r warning=FALSE}
pdf("./Figures/Alpha Diversity by plant zone 6.pdf", width = 8, height = 5 )
physeq.plot.richness.plant
dev.off()
```

```{r}
png("./Figures/WHO_Alpha_Diversity_Rice_zone.png", units="in", width = 8, height = 5, res = 1000)
physeq.plot.richness.plant
dev.off()
```

