---
title: "Venn"
output: html_document
date: "2024-01-14"
---

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(VennDiagram)
library(car)
#library(Hmisc)
```


```{r}
otus <- read.table("DroughtITS.final_fix.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP4: Import taxonomy table for analysis using Phyloseq
This taxonomy table was created from STEP2.
```{r}
taxmat <- read.table("DroughtITS.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP2: Import tree file for analysis using Phyloseq

```{r warning=FALSE}
treefile = "DroughtITS.tree.phy"
tree = read.tree(treefile)
```

###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r}
meta = read.table("DroughtITS.mapping_file.fix.final.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

Check if your metadata file has been import successfully and correctly, the output will show a table of your metadata file (mapping file). *If you do not have header, you might start your first row with # (remove # and reload your mapping file).

```{r}
head(meta)
```

Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```


###STEP4: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
physeq
```

```{r}
physeq = subset_taxa(physeq, Kingdom == "Fungi")
#physeq = subset_samples(physeq)

```

```{r}
sampleData <- sample_data(meta)
```


```{r}
physeq <- phyloseq(OTU,TAX,sampleData)
physeq
physeq <- subset_taxa(physeq, Kingdom == "Fungi")
physeq
```


```{r}
physeq.prune = prune_taxa(taxa_sums(physeq) > 1, physeq)
physeq.prune
```
```{r}
phyloseq::tax_table(physeq.prune)
```

```{r}
physeq.zone = subset_samples(physeq.prune)
```

##Make taxonomy table into a matrix and relabel NA as unknown
```{r}
tax.fun <- as(tax_table(physeq.zone),"matrix")
head(tax.fun)
tax.fun[is.na(tax.fun)] <- "Unknown"
head(tax.fun)
```

###Convert tax table back to phyloseq object and generate phyloseq object with new tax table
```{r}
TAX.fun <- tax_table(tax.fun)
fun.3 <- phyloseq(sample_data(physeq.zone),otu_table(physeq.zone),TAX.fun)
fun.3
```


```{r}
glom.fun <- speedyseq::tax_glom(fun.3,taxrank = "Genus")
glom.fun
```

```{r}
head(tax_table(glom.fun))
```


##plant##

##Transform OTU table to show relative abundance
##Samples can also be merged together by a variable in the mapping file

```{r}

fun.abund <- merge_samples(glom.fun, "plant")
sample_data(fun.abund)$plant <- factor(sample_names(fun.abund))
fun.abund = transform_sample_counts(fun.abund, function(x) x / sum(x))
fun.abund

```


```{r}
data_glom.fun <- speedyseq::psmelt(fun.abund)
data_glom.fun$Genus <- as.character(data_glom.fun$Genus)
```

##If a phylum has less than 1% abundance, phylum name is changed to <1% abund.

```{r}
data_glom.fun$Genus[data_glom.fun$Abundance < 0.01] <- "<0.1% abund."
```

Count the levels present in the Phylum column

```{r}
Count = length(unique(data_glom.fun$Genus))
Count
```

Print out unique phyla names for insertion into barplots in next step.

```{r}
unique((data_glom.fun$Genus))
```

```{r}
venn_data <- list(Rice = unique(data_glom.fun$Genus[data_glom.fun$plant == "Rice"]),
                  Sugarcane = unique(data_glom.fun$Genus[data_glom.fun$plant == "Sugarcane"]),
                  Rubber_tree = unique(data_glom.fun$Genus[data_glom.fun$plant == "Rubber tree"]),
                  Cassava = unique(data_glom.fun$Genus[data_glom.fun$plant == "Cassava"]))
```




```{r}
number_of_taxa <- data_glom.fun$Genus %>% 
  filter(Abundance > 0) %>% 
  group_by(plant, Genus) %>% 
  summarise(n = length(Abundance))
```


```{r}
library(ggvenn)
```


```{r}
venn <- ggvenn(venn_data, 
       fill_color = c("#fde725", "#ed6925", "#9c179e", "#0d0887"),
       fill_alpha = 0.5,
       show_percentage = FALSE,
       text_size = 5,
       set_name_size = 5)
venn
```

```{r}
png("./Figures/venn.png",units = "in", width = 7.5, height = 6, res = 600 )
venn
dev.off()
```





