---
title: "Indicator species"
output: html_document
date: "2023-04-20"
---

```{r warning=FALSE, message=FALSE}
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(data.table)
library(tidyr)
library(tidyverse)
library(multcompView)
library(VennDiagram)
library(car)
library(microViz)
#library(Hmisc)
```


```{r}
otus <- read.table("./Data/DroughtITS.final_fix.txt",header=T,sep="\t",row.names=1)
otumat <- as(as.matrix(otus), "matrix")
OTU = otu_table(otumat, taxa_are_rows = TRUE)
```

###STEP4: Import taxonomy table for analysis using Phyloseq
This taxonomy table was created from STEP2.
```{r}
taxmat <- read.table("./Data/DroughtITS.taxonomy.fix.txt", header=T,sep="\t",row.names=1)
taxmat <- as(as.matrix(taxmat),"matrix")
TAX = tax_table(taxmat)
```

###STEP2: Import tree file for analysis using Phyloseq

```{r warning=FALSE}
treefile = "./Data/DroughtITS.tree.phy"
tree = read.tree(treefile)
```

###STEP3: Import mapping file for analysis using Phyloseq

1.Check mapping file before import to R, R will automatically change sample name that starts with number or contain “-” in sample name. If you get error in this step, you should check sample name first. 

2.First column of first row should not start with #, R will not read the first row that starts with #

3. You can choose which samples to include in analysis by indicating specific group in the column

```{r}
meta = read.table("./Data/DroughtITS.mapping_file.fix.txt",header=TRUE,row.names=1,sep="\t",stringsAsFactors=FALSE)
```

Check if your metadata file has been import successfully and correctly, the output will show a table of your metadata file (mapping file). *If you do not have header, you might start your first row with # (remove # and reload your mapping file).

```{r}
head(meta)
```

Construct sample_data-class using imported metadata
```{r warning=FALSE}
sampleData <- sample_data(meta)
```


###STEP4: Construct Phyloseq object
To construct phyloseq object, otu table, taxonomy table, and sampleData are required. Phylogenetic tree can be included, but it is not necessary for constructing phyloseq object.
Construct Phyloseq object called "Physeq"

```{r warning=FALSE}
physeq = phyloseq(OTU,TAX,sampleData)
```

###STEP8: Check phyloseq object
This should indicate that your physeq is a "phyloseq-class experiment-level object"

```{r}
physeq
```

```{r}
#critical group #non of the critical fungi was detected
#physeq.1 = subset_taxa(physeq, Species == "Cryptococcus neoformans")
#physeq.2 = subset_taxa(physeq, Species == "Candida auris")
#physeq.3 = subset_taxa(physeq, Species == "Aspergillus fumigatus")
#physeq.4 = subset_taxa(physeq, Species == "Candida albican")

#High group #only candida tropicalis
#physeq.5 = subset_taxa(physeq, Species == "Nakaseomyces glabrata")
#physeq.6 = subset_taxa(physeq, Genus == "Histoplasma")
physeq.7 = subset_taxa(physeq, Species == "Candida tropicalis")
#physeq.8 = subset_taxa(physeq, Species == "Candida parapsilosis")

#physeq.17 = subset_taxa(physeq, Genus == "Madurella")
physeq.18 = subset_taxa(physeq, Species == "Falciformispora senegalensis")
physeq.19 = subset_taxa(physeq, Species == "Curvularia lunata")
#physeq.20 = subset_taxa(physeq, Species == "Zopfia rosatii")
physeq.21 = subset_taxa(physeq, Genus == "Acremonium")
physeq.22 = subset_taxa(physeq, Genus == "Fusarium")
physeq.23 = subset_taxa(physeq, Genus == "Rhizopus")
physeq.24 = subset_taxa(physeq, Genus == "Mucor")
physeq.25 = subset_taxa(physeq, Genus == "Lichtheimia")

#Medium group
physeq.9 = subset_taxa(physeq, Genus == "Scedosporium")
#physeq.10 = subset_taxa(physeq, Species == "Lomentospora prolificans")
#physeq.11 = subset_taxa(physeq, Genus == "Coccidioides")
#physeq.12 = subset_taxa(physeq, Species == "Pichia kudriavzeveii")
#physeq.13 = subset_taxa(physeq, Species == "Cryptococcus gattii")
physeq.14 = subset_taxa(physeq, Species == "Talaromyces marneffei")
#physeq.15 = subset_taxa(physeq, Species == "Pneumocystis jirovecii")
#physeq.16 = subset_taxa(physeq, Genus == "Paracoccidioides")


# High group: Candida tropicalis and Medium group: Falciformispora senegalensis, Curvularia lunata, Acremonium spp, Fusarium spp, Rhizopus spp, Mucor spp, Lichtheimia, Scedosporium, Talaromyces marneffei

physeq.merge = merge_phyloseq(physeq.7, physeq.9, physeq.14, physeq.18, physeq.19, physeq.21, physeq.22, physeq.23, physeq.24, physeq.25)

physeq.merge
```
Convert phyloseq OTU table to reflect species

Glomerate at Species level (However, not all of species will be unique)
See for ref: https://github.com/joey711/phyloseq/issues/927

```{r}
glom.fun <- speedyseq::tax_glom(physeq.merge,taxrank = "Species")
glom.fun
```

```{r}
glomTax = tax_table(glom.fun)[,"Species"]
```

Combine OTU table and Tax table
```{r}
glomOTU = otu_table(glom.fun)
glomTable = merge(glomOTU,glomTax,by=0,all=TRUE)
head(glomTable)
```

Use Speceis as rowname and remove prior rowname and species column
```{r}
rownames(glomTable) = glomTable[,"Species"]
glomTable$Row.names = NULL
glomTable$Species = NULL
```


Transpose Rows and columns for desired labdsv format
```{r}
ASV = t(glomTable)
```

Remove columns with 0 species counts

```{r}
ASV.no0 = ASV[, colSums(ASV != 0) > 0]
```

Subset sampleData phyloseq object for the desired variable (e.g. treatment, etc.)
```{r}
ASV_provinces = subset(sampleData,select = "province")
```

Make the varible numeric
```{r}
ASV_provinces$province
```

```{r}
ASV_provinces$province = as.numeric(as.factor(ASV_provinces$province))
ASV_provinces$province
```

```{r}
class(ASV_province$province)
```

Run indval code from lab.dsv 
```{r}
ThaiDrought.indic.province = indval(ASV.no0,ASV_province$province)
```

```{r}
summary(ThaiDrought.indic.province, p = 0.05, type = "short")
```



